The lattice of acyclic pipe dreams

### 2022-10-17 acyclicFacets/draft/latticePipeDreams.tex
cleaning Section 2

Two pipe dreams are related by a \defn{flip} if they define the same permutation of the pipes and only differ by the position of a cross and an elbow.
An elbow~$e$ in a pipe dream~$P$ is \defn{flippable} if the two pipes passing through elbow~$e$ have a crossing~$c$, and the flip exchanges the elbow~$e$ with the cross~$c$.
See \cref{fig:pipeDreams}.
The flip is \defn{increasing} if the elbow~$e$ is weakly southwest of the crossing~$c$.
For example, the flip of \cref{fig:pipeDreams} is increasing from left to right.
%
The \defn{increasing flip graph} is the graph of increasing flips on~$\pipeDreams(\omega)$.
It is clearly a directed acyclic graph, and it has a unique source and a unique sink \cite{PilaudPocchiola}, called the \defn{greedy} and \defn{antigreedy} pipe dreams, and denoted~$\greedyPipeDream$ and~$\antiGreedyPipeDream$.
\vincent{In \cite{PilaudStump-ELlabeling}, these are called positive and negative greedy facets. Also the notations are different...}

@@

@@

\subsection{Reversing pipe dreams}
\label{subsec:reversingPipeDreams}

We say that a pipe dream~$P \in \pipeDreams_n$ is \defn{reversing} if it fixes the first pipe and reverses the order of the remaining pipes, \ie if~$\omega_P = [1, n, n-1, \dots, 2]$ (in one-line notation).
We denote by~$\reversingPipeDreams_n$ the reversing pipe dreams of~$\pipeDreams_n$.
As observed by different authors~\cite{Woo, PilaudPocchiola, Pilaud-these, Stump}, the pipe dreams of~$\reversingPipeDreams_n$ belong to the Catalan family.
\cref{fig:bijection} illustrates explicit bijections between the pipe dreams of~$\reversingPipeDreams_n$, the binary trees on $n$ vertices, and the triangulations of a convex $(n+2)$-gon.
\vincent{figure starts with pipe $0$}

\begin{figure}[h]
	\centerline{\includegraphics[scale=1.3]{dualiteTriangulation}}
	\caption{The bijection between reversing pipe dreams (left), binary trees (middle) and triangulations (right).}
	\label{fig:bijection}
\end{figure}

More precisely, the map which sends an elbow in row~$i$ and column~$j$ of the triangular shape to the diagonal~$[i,n+1-j]$ of the~$(n+2)$-gon provides the following correspondence:
\begin{center}
\begin{tabular}{r@{$\quad\longleftrightarrow\quad$}l}
pipe dream~$P \in \reversingPipeDreams_n$ & triangulation~$P\duality$ of the $(n+2)$-gon, \\
$i$th pipe of~$P$ & $i$th triangle of~$P\duality$ (with central vertex~$i$), \\
elbows of~$P$ & diagonals of~$P\duality$, \\
crosses of~$P$ & common bisectors between triangles of~$P\duality$, \\
contact graph of~$P$ & dual binary tree of~$P\duality$, \\
elbow flips in~$P$ & diagonal flips in~$P\duality$.
\end{tabular}
\end{center}

\begin{remark}
All reversing pipe dreams are acyclic.
Indeed, there contact graphs are (oriented) binary trees, which are certainly acyclic.
\end{remark}

\subsection{Elementary properties of pipe dreams}
\label{subsec:elementaryProperties}

We gather in this section some elementary properties of pipe dreams needed later.
\vincent{Not sure we use all these. At least say that they are just technical points.}

\begin{lemma}
\label{lem:horiVertCrossings}
For any pipe dream~$P \in \pipeDreams(\omega)$, the pipe~$j$ of~$P$ (which enters at row~$j$ and exits at column~$\omega^{-1}(j)$) crosses
\begin{itemize}
\item vertically the pipes~$i$ such that~$i < j$ while~$\omega^{-1}(i) > \omega^{-1}(j)$,
\item horizontally the pipes~$k$ such that~$j < k$ while~$\omega^{-1}(j) > \omega^{-1}(k)$.
\end{itemize}
\end{lemma}

\begin{proof}
For~$i < j$, if~$\omega^{-1}(i) > \omega^{-1}(j)$, then the pipes~$i$ and~$j$ have to cross exactly once, while if~$\omega^{-1}(i) < \omega^{-1}(j)$ the pipes~$i$ and~$j$ cannot cross.
The same argument applies for~${k > j}$.
\end{proof}

\begin{lemma}
\label{lem:countingElbows}
For any pipe dream~$P \in \pipeDreams(\omega)$, the pipe~$j$ (which enters at row~$j$ and exits at column~$\omega^{-1}(j)$) has precisely
\begin{itemize}
\item $\noninversions{\omega}{j}$ many southeast elbows\!\SEelbow{}
\item $1+\noninversions{\omega}{j}$ many northwest elbows~\WNelbow{}
\item $j-1-\noninversions{\omega}{j}$ vertical crossings~\NScross
\item $\omega^{-1}(j)-1-\noninversions{\omega}{j}$ horizontal crossings~\WEcross
\end{itemize}
where~$\noninversions{\omega}{j} \eqdef \#\set{i \in [n]}{i < j \text{ and } \omega^{-1}(i) < \omega^{-1}(j)}$.
\end{lemma}

@@

@@

Note that the reciprocal assertion of \cref{lem:rectangle} is false.
We conclude by two consequences of \cref{lem:rectangle}.

@@

\begin{lemma}
\label{lem:consequenceRectangle2}
Let~$P$ be a pipe dream and let~$i,j,k$ be three pipes of~$P$ such that~$\min(i,k) < j$ and~$\min(\omega^{-1}(i), \omega^{-1}(k)) < \omega^{-1}(j)$.
If~$i \to k$ in~$P\contact$, then either~$i \contactLess{P} j \contactMore{P} k$ or~$i \contactMore{P} j \contactLess{P} k$.
\vincent{not sure we really use this version}
\end{lemma}

\begin{proof}
Let~$e$ denote the contact of~$i$ and~$k$ in~$P$.
Decompose the triangular shape into three regions: the region~$A$ of all points located southwest of~$e$, the region~$B$ of all points located northwest or southeast of~$e$, and the region~$C$ of all points located northeast of~$e$.
Since~$\min(i,k) < j$ and~$\min(\omega^{-1}(i), \omega^{-1}(k)) < \omega^{-1}(j)$, the pipe~$j$ starts in region~$A$ and ends in region~$C$.
Hence, the pipe~$j$ has an elbow~$f$ in region~$B$.
We thus obtain that~$i \contactLess{P} j \contactMore{P} k$ if this $f$ is southeast of~$e$, and that~$i \contactMore{P} j \contactLess{P} k$ if $f$ is northwest of~$e$.
\end{proof}

@@

\vincent{Mention here and throughout the classical sylvester example}

@@

We will see in \cref{sec:algorithms} different algorithms to compute the pipe dream~$P \in \acyclicPipeDreams(\omega)$ of which a given permutation~$\pi \in [e,\omega]$ is a linear extension.
These algorithms are however not needed for the proof of \cref{thm:partitionPipeDreams}, which we break into the next three elementary lemmas.

@@

We prove this statement by checking both conditions of \cref{prop:characterizationCongruences}.
For the first condition, we need the following classical characterization of weak order intervals, see \cite{BjornerWachs} or~\cite{ChatelPilaudPons}.

\begin{proposition}[{\cite[Thm.~6.8]{BjornerWachs}}]
\label{prop:WOIP}
The set~$\linearExtensions(\less)$ of linear extensions of a poset~$\less$ on~$[n]$ forms an interval~$I$ of the weak order if and only if for every~$i < j < k$,
\[
i \less k \implies i \less j \text{ or } j \less k
\qquad\text{and}\qquad
i \more k \implies i \more j \text{ or } j \more k.
\]
Moreover, the inversions of~$\min(I)$ are the pairs~$i,j \in [n]$ with $i < j$ and $i \more j$, and the non-inversions of~$\max(I)$ are the pairs~$i,j \in [n]$ with $i < j$ and $i \less j$.
\vincent{Not sure I want to use that last sentence later.}
\end{proposition}

@@

For a congruence~$\equiv$ of a lattice~$L$, the \defn{lattice quotient}~$L/{\equiv}$ is the lattice on the classes of~$\equiv$ where for any two congruence classes~$X$ and~$Y$, 
\begin{itemize}
\item $X \le Y$ in~$L/{\equiv}$ if and only if there exist representatives~$x \in X$ and~$y \in Y$ such that~$x \le y$ in~$L$, or equivalently $\min(X) \le \min(Y)$, or equivalently $\max(X) \le \max(Y)$,
\item $X \meet Y$ (resp.~$X \join Y$) is the congruence classe of~$x \meet y$ (resp.~of~$x \join y$) for arbitrary representatives~$x \in X$ and~$y \in Y$.
\end{itemize}
In this section, we prove the following statement.

\begin{theorem}
\label{thm:pipeDreamQuotient}
The lattice quotient~$[e,\omega]/{\equiv_\omega}$ is isomorphic to the increasing flip poset on~$\acyclicPipeDreams(\omega)$.
\end{theorem}

To prove this result, we need the following auxiliary statement.

@@

\begin{proof}
Say that~$i < j$ while~$\omega^{-1}(i) > \omega^{-1}(j)$ and that~$i \to j$ is an arc of~$P\contact$ while~$j \to i$ is an arc of~$P'{}\contact$.
Since $P\contact$ is acyclic, there is no path from~$j$ to~$i$ in~$P\contact$.
Assume by means of contradiction that there is a path~$i \to k_1 \to \dots \to k_p \to j$ in~$P\contact$ with~$p \ge 1$.
Since the arcs of~$P'{}\contact$ are the arcs of~$P\contact$ where only extremities~$i$ and~$j$ can be changed, $P'{}\contact$ contains the path~$k_1 \to \dots \to k_p$ and at least one of the arcs~$i \to k_1$ or~$j \to k_1$, and at least one of the arcs~$k_p \to j$ or~$k_p \to i$.
Consequently, since~$P'{}\contact$ contains the arc~$j \to i$ and is acyclic, it must contain the path~$j \to k_1 \to \dots \to k_p \to i$.
We thus obtained that~$i \contactLess{P} k_1 \contactLess{P} j$ while~$i \contactMore{P'} k_1 \contactMore{P'} j$, and~$k_1$ has a contact with~$i$ in~$P$ and with~$j$ in~$P'$.

Consider now the elbow~$e$ of~$P$ which is a crossing in~$P'$ and the elbow~$e'$ of~$P'$ which is a crossing of~$P$.
Let~$R$ the rectangle with corners~$e$ and~$e'$.
Since~$k_1$ has a contact with~$i$ in~$P$ and with~$j$ in~$P'$, it must pass inside~$R$.
Since~$i \contactLess{P} k_1 \contactLess{P} j$ and~$i \contactMore{P'} k_1 \contactMore{P'} j$, the pipe~$k$ has no elbow located northwest or southeast of~$e$ or~$e'$, hence no elbow located north, south, west or east of~$R$.
We thus obtain that~$k$ must be straight before it reaches~$R$, and after it leaves~$R$.
Hence $k_1 < i < j$ and~$\omega^{-1}(k_1) > \omega^{-1}(j) > \omega^{-1}(i)$.
By \cref{lem:consequenceRectangle1}, this contradicts~$k_1 \contactLess{P} j$~and~$i \contactMore{P'} k_1$.
\end{proof}

\begin{proof}[Proof of \cref{thm:pipeDreamQuotient}]
It suffices to prove that the following conditions are equivalent for two distinct pipe dreams~$P,P' \in \acyclicPipeDreams(\omega)$:
\begin{enumerate}[(i)]
\item $P$ covers~$P'$ in the increasing flip poset on~$\acyclicPipeDreams(\omega)$,
\item there exist linear extensions~$\pi$ of~$P$ and~$\pi'$ of~$P'$ such that~$\pi$ covers~$\pi'$ in weak order.
\end{enumerate}
\para{(ii) $\Rightarrow$ (i)}
Follows from \cref{lem:lowerSetPipeDreams}.

\para{(i) $\Rightarrow$ (ii)}
Assume that~$P$ covers~$P'$ in the increasing flip poset.
Let~$i < j$ be the two pipes involved in the flip between~$P$ and~$P'$.
Hence, $j \to i$ is an arc of~$P$ while~$i \to j$ is an arc of~$P'$.
By \cref{lem:flippable}, there is no directed path from~$j$ to~$i$ in~$P$ besides the arcs~$j \to i$ (there might be more than one such arc).
Hence, there exists a linear extension~$\pi$ of~$P$ where~$j$ and~$i$ are consecutive.
Write~$\pi \eqdef UjiV$ and define~$\pi' \eqdef UijV$.
Since~$j \to i$ is an arc of~$P$, $\pi'$ is not a linear extension of~$P$.
Hence, by \cref{lem:lowerSetPipeDreams}, $\pi'$ is a linear extension of~$P'$.
\end{proof}

@@

\begin{proposition}
For any pipe dreams~$P, P' \in \acyclicPipeDreams(\omega)$, the following assertions are equivalent:
\begin{enumerate}[(i)]
\item $P < P'$ in the increasing flip poset on~$\acyclicPipeDreams(\omega)$,
\item there exist linear extensions~$\pi$ of~$P$ and~$\pi'$ of~$P'$ such that~$\pi < \pi'$ in weak order,
\item the minimal (resp.~max) linear extensions~$\pi$ of~$P$ and~$\pi'$ of~$P$ satisfy~$\pi < \pi'$ in weak order.
\item there is no~$i < j$ such that~$i \contactMore{P} j$ and~$i \contactLess{P'} j$,
\item for all~$i < j$, if~$i \contactMore{P} j$, then~$j \contactMore{P'} i$,
\item for all~$i < j$, if~$i \contactLess{P'} j$, then~$i \contactLess{P} j$,
\end{enumerate}
\end{proposition}

\begin{proof}
We already proved that (i) $\Leftrightarrow$ (ii).
\vincent{todo}
\end{proof}

@@

@@

>>>>

Two pipe dreams are related by a \defn{flip} if they define the same permutation of the pipes and only differ by the position of a cross and an elbow.
An elbow~$e$ in a pipe dream~$P$ is \defn{flippable} if the two pipes passing through elbow~$e$ have a crossing~$c$, and the flip exchanges the elbow~$e$ with the cross~$c$.
See \cref{fig:pipeDreams}.
The flip is \defn{increasing} if the elbow~$e$ is weakly southwest of the crossing~$c$.
For example, the flip of \cref{fig:pipeDreams} is increasing from left to right.
%
The \defn{increasing flip graph} is the graph of increasing flips on~$\pipeDreams(\omega)$.
It is clearly a directed acyclic graph, and it has a unique source and a unique sink \cite{PilaudPocchiola}, called the \defn{greedy} and \defn{antigreedy} pipe dreams, and denoted~$\greedyPipeDream$ and~$\antiGreedyPipeDream$.
\vincent{In \cite{PilaudStump-ELlabeling}, these are called positive and negative greedy facets. Also the notations are different...}
The \defn{increasing flip poset} is the reflexive and transitive closure of the increasing flip graph on~$\pipeDreams(\omega)$.

@@

\begin{example}
\label{exm:Tamari1}
We say that a pipe dream is \defn{reversing} if it fixes the first and last pipes and reverses the order of the remaining pipes.
In this case, it is natural to label the pipes from~$0$ to~$n+1$, so that the permutation of the pipes is~$\rho_n \eqdef [0, n, n-1, \dots, 2, 1, n+1]$.
% We denote by~$\reversingPipeDreams_n$ the reversing pipe dreams of~$\pipeDreams_n$.
As observed in~\cite{Woo, PilaudPocchiola, Pilaud-these, Stump}, the family of reversing pipe dreams belong to the Catalan families.
\cref{fig:bijection} illustrates explicit bijections between reversing pipe dreams of~$\pipeDreams(\rho_n)$, binary trees with $n$ internal nodes, and the triangulations of a convex $(n+2)$-gon.
More precisely, the map which sends an elbow in row~$i$ and column~$j$ of the triangular shape to the diagonal~$[i,n+1-j]$ of the~$(n+2)$-gon provides the following correspondence: \\[.3cm]
\centerline{
\begin{tabular}{r@{$\quad\longleftrightarrow\quad$}l}
pipe dream~$P \in \reversingPipeDreams_n$ & triangulation~$P\duality$ of the $(n+2)$-gon, \\
pipe $i$ of~$P$ & triangle $i\duality$ of~$P\duality$ (with central vertex~$i$), \\
elbow between pipes~$i$ and~$j$ of~$P$ & common side of triangles~$i\duality$ and~$j\duality$ of~$P\duality$, \\
cross between pipes~$i$ and~$j$ of~$P$ & common bisector between triangles~$i\duality$ and~$j\duality$ of~$P\duality$, \\
contact graph of~$P$ & dual binary tree of~$P\duality$, \\
elbow flips in~$P$ & diagonal flips in~$P\duality$. \\
%increasing flip poset & Tamari lattice.
\end{tabular}} \\[.3cm]
%
\begin{figure}
	\centerline{\includegraphics[scale=1.3]{dualiteTriangulation}}
	\caption{The bijection between reversing pipe dreams (left), binary trees (middle) and triangulations (right).}
	\label{fig:bijection}
\end{figure}
%
As a consequence, this map sends the increasing flip poset on reversing pipe dreams to the \defn{Tamari lattice} on binary trees.
This lattice is defined as the transitive closure of right rotations on binary trees, and is obtained as the quotient of the weak order by the \defn{sylvester congruence}.
\vincent{Not sure where we want that...}
An important point here is that all reversing pipe dreams are acyclic, since there contact graphs are (oriented) binary trees.
\end{example}

@@

\subsection{Crossing and contact properties}
\label{subsec:crossingsContacts}

We now gather some elementary properties of crossings and contacts that will be needed later to construct pipe dreams from permutations.
For a pipe dream~$P \in \pipeDreams(\omega)$, we call pipe~$j$ the pipe which enters at row~$j$ and exits at column~$\omega^{-1}(j)$.

\begin{lemma}
\label{lem:horiVertCrossings}
For any pipe dream~$P \in \pipeDreams(\omega)$, the pipe~$j$ of~$P$ crosses precisely
\begin{itemize}
\item vertically the pipes~$i$ such that~$i < j$ while~$\omega^{-1}(i) > \omega^{-1}(j)$,
\item horizontally the pipes~$k$ such that~$j < k$ while~$\omega^{-1}(j) > \omega^{-1}(k)$.
\end{itemize}
\end{lemma}

\begin{proof}
For~$i < j$, if~$\omega^{-1}(i) > \omega^{-1}(j)$, then the pipes~$i$ and~$j$ have to cross exactly once (and $j$ must be the vertical pipe at that crossing), while if~$\omega^{-1}(i) < \omega^{-1}(j)$ the pipes~$i$ and~$j$ cannot cross.
The same argument applies for~${k > j}$.
\end{proof}

\begin{lemma}
\label{lem:countingElbows}
For any pipe dream~$P \in \pipeDreams(\omega)$, the pipe~$j$ has precisely
\begin{itemize}
\item $\noninversions{\omega}{j}$ many southeast elbows\!\SEelbow{}
\item $1+\noninversions{\omega}{j}$ many northwest elbows~\WNelbow{}
\item $j-1-\noninversions{\omega}{j}$ vertical crossings~\NScross
\item $\omega^{-1}(j)-1-\noninversions{\omega}{j}$ horizontal crossings~\WEcross
\end{itemize}
where~$\noninversions{\omega}{j} \eqdef \#\set{i \in [n]}{i < j \text{ and } \omega^{-1}(i) < \omega^{-1}(j)}$.
\end{lemma}

@@

\subsection{Contact graph properties}
\label{subsec:contactGraph}

We now state a simple observation about the poset~$\contactLess{P}$, and two of its consequences, that will play essential roles in the proofs in \cref{sec:latticeAcyclicPipeDreams}.

@@

Note that the reciprocal assertion of \cref{lem:rectangle} is false.
We conclude this section by two consequences of \cref{lem:rectangle}.

@@

\begin{lemma}
\label{lem:consequenceRectangle2}
Let~$P$ be a pipe dream and let~$i,j,k$ be three pipes of~$P$ such that~$i < j < k$ and~$\omega^{-1}(i) > \omega^{-1}(j) > \omega^{-1}(k)$.
If~$i \to k$ in~$P\contact$, then either~$i \contactLess{P} j \contactMore{P} k$ or~$i \contactMore{P} j \contactLess{P} k$.
\end{lemma}

\begin{proof}
Let~$e$ denote the contact of pipes~$i$ and~$k$ in~$P$.
Decompose the triangular shape into three regions: the region~$A$ of all points located southwest of~$e$, the region~$B$ of all points located northwest or southeast of~$e$, and the region~$C$ of all points located northeast of~$e$.
Since~$i < j$ and~$\omega^{-1}(j) > \omega^{-1}(k)$, the pipe~$j$ starts in region~$A$ and ends in region~$C$.
Hence, the pipe~$j$ has an elbow~$f$ in region~$B$.
We thus obtain that~$i \contactLess{P} j \contactMore{P} k$ if this $f$ is southeast of~$e$, and that~$i \contactMore{P} j \contactLess{P} k$ if $f$ is northwest of~$e$.
\end{proof}

%\begin{lemma}
%\label{lem:consequenceRectangle2}
%Let~$P$ be a pipe dream and let~$i,j,k$ be three pipes of~$P$ such that~$\min(i,k) < j$ and~$\min(\omega^{-1}(i), \omega^{-1}(k)) < \omega^{-1}(j)$.
%If~$i \to k$ in~$P\contact$, then either~$i \contactLess{P} j \contactMore{P} k$ or~$i \contactMore{P} j \contactLess{P} k$.
%\vincent{not sure we really use this extended version}
%\end{lemma}
%
%\begin{proof}
%Let~$e$ denote the contact of pipes~$i$ and~$k$ in~$P$.
%Decompose the triangular shape into three regions: the region~$A$ of all points located southwest of~$e$, the region~$B$ of all points located northwest or southeast of~$e$, and the region~$C$ of all points located northeast of~$e$.
%Since~$\min(i,k) < j$ and~$\min(\omega^{-1}(i), \omega^{-1}(k)) < \omega^{-1}(j)$, the pipe~$j$ starts in region~$A$ and ends in region~$C$.
%Hence, the pipe~$j$ has an elbow~$f$ in region~$B$.
%We thus obtain that~$i \contactLess{P} j \contactMore{P} k$ if this $f$ is southeast of~$e$, and that~$i \contactMore{P} j \contactLess{P} k$ if $f$ is northwest of~$e$.
%\end{proof}

@@

As already mentioned, the increasing flip poset on reversing pipe dreams is isomorphic to the Tamari lattice, which is a well known lattice quotient of the weak order.
In this section, we extend this result to any permutation~$\omega$, by proving that the transitive closure of the increasing flip graph on acyclic pipe dreams of~$\acyclicPipeDreams(\omega)$ is the Hasse diagram of a lattice quotient of the weak order interval~$[e,\omega]$.

@@

\begin{example}
\label{exm:Tamari2}
Following \cref{exm:Tamari1}, observe that the permutations of~$\{0, 1, \dots, n, n+1\}$ below $\rho_n$ are precisely the permutations of the form~$[0, \pi, n+1]$ for some~$\pi \in \fS_n$.
It is well-known that any permutation~$\pi \in \fS_n$ is a linear extension of a unique binary tree.
This binary tree can be obtained by inserting~$\pi$ from right to left in a binary search tree.
Hence any permutation of~$\{0, 1, \dots, n, n+1\}$ below~$\rho_n$ is a linear extension of a unique reversing pipe dream on~$n+2$ pipes.
For instance, the pipe dream of \cref{fig:bijection} has linear extensions~$0421356$, $0423156$, $0421536$, $0423516$, $0425136$, $0425316$, $0452136$, $0452316$.
\end{example}

We will see in \cref{sec:algorithms} insertion algorithms to compute the pipe dream~$P \in \acyclicPipeDreams(\omega)$ of which a given permutation~$\pi \in [e,\omega]$ is a linear extension, for any permutation~$\omega$.
These algorithms are however not needed for the proof of \cref{thm:partitionPipeDreams}, which we break into the following three lemmas.

@@

\begin{example}
\label{exm:Tamari3}
Following \cref{exm:Tamari1,exm:Tamari2}, observe that the congruence~$\equiv_{\rho_n}$ on the permutations of~$\{0, 1, \dots, n, n+1\}$ below~$\rho_n$ corresponds to the sylvester congruence on~$\fS_n$.
The classes of this congruence are the sets of linear extensions of the binary trees (considered as posets, labeled in inorder, and oriented towards their roots).
It can also be defined by the classical rewriting rule~$UikVjW \equiv UkiVjW$ for~$i < j < k$.
\end{example}

We will discuss in \cref{subsec:rewritingAlgorithm} rewriting rules for the pipe dream congruence~$\equiv_\omega$, for any permutation~$\omega$.
This rewriting rule is however not needed for the proof of \cref{thm:pipeDreamCongruence}.
We prove it by checking both conditions of \cref{prop:characterizationCongruences}.
For the first condition, we need the following classical characterization of weak order intervals, see \cite{BjornerWachs} or~\cite{ChatelPilaudPons}.

\begin{proposition}[{\cite[Thm.~6.8]{BjornerWachs}}]
\label{prop:WOIP}
The set~$\linearExtensions(\less)$ of linear extensions of a poset~$\less$ on~$[n]$ forms an interval~$I$ of the weak order if and only if for every~$i < j < k$,
\[
i \less k \implies i \less j \text{ or } j \less k
\qquad\text{and}\qquad
i \more k \implies i \more j \text{ or } j \more k.
\]
Moreover, the inversions of~$\min(I)$ are the pairs~$i,j \in [n]$ with $i < j$ and $i \more j$, and the non-inversions of~$\max(I)$ are the pairs~$i,j \in [n]$ with $i < j$ and $i \less j$.
\end{proposition}

@@

For a congruence~$\equiv$ of a lattice~$L$, the \defn{lattice quotient}~$L/{\equiv}$ is the lattice on the classes of~$\equiv$ where for any two congruence classes~$X$ and~$Y$, 
\begin{itemize}
\item $X \le Y$ in~$L/{\equiv}$ if and only if there exist representatives~$x \in X$ and~$y \in Y$ such that~$x \le y$ in~$L$, or equivalently $\min(X) \le \min(Y)$, or equivalently $\max(X) \le \max(Y)$,
\item $X \meet Y$ (resp.~$X \join Y$) is the congruence classe of~$x \meet y$ (resp.~of~$x \join y$) for arbitrary representatives~$x \in X$ and~$y \in Y$.
\end{itemize}
In this section, we aim at the following statement.

\begin{theorem}
\label{thm:pipeDreamQuotient}
The increasing flip graph on~$\acyclicPipeDreams(\omega)$ is isomorphic to the Hasse diagram of the lattice quotient~$[e,\omega]/{\equiv_\omega}$.
\end{theorem}

\begin{example}
\label{exm:Tamari4}
Following \cref{exm:Tamari1,exm:Tamari2,exm:Tamari3}, observe that the increasing flip poset on reversing pipe dreams is isomorphic to the Tamari lattice, which is the quotient of the weak order by the sylvester congruence.
\end{example}

\begin{remark}
\label{rem:pipeDreamQuotient}
The increasing flip poset on~$\pipeDreams(\omega)$ is the transitive closure of the increasing flip graph on~$\pipeDreams(\omega)$.
Observe that the two natural ways to restrict to the acyclic pipe dreams of~$\acyclicPipeDreams(\omega)$ (restrict either the flip graph or the flip poset) do not coincide in general.
Namely, the transitive closure of the subgraph of the increasing flip graph induced by~$\acyclicPipeDreams(\omega)$ may have strictly less relations than the subposet of the increasing flip poset induced by~$\acyclicPipeDreams(\omega)$.
\cref{fig:counterExampleRestrictionIncreasingFlipPoset} illustrates two acyclic pipe dreams of~$\acyclicPipeDreams(126543)$ connected by a sequence of increasing flips in~$\pipeDreams(126543)$ but by no sequence of increasing flips in~$\acyclicPipeDreams(126543)$.
%	* type A5,
%	* Q = [1, 2, 3, 4, 5, 1, 2, 3, 4, 1, 2, 3, 1, 2, 1] 
%	* w = [1, 2, 3, 1, 2, 1]
%	* flips: 
%	(0, 2, 3, 4, 5, 6, 7, 8, 12)
%	(2, 3, 4, 5, 6, 7, 8, 10, 12) 
%	(3, 4, 5, 6, 7, 8, 10, 11, 12)
%	(3, 4, 5, 7, 8, 10, 11, 12, 14)

\begin{figure}[ht]
	\centerline{\includegraphics[scale=1.3]{counterExampleRestrictionIncreasingFlipPoset}}
	\caption{Two acyclic pipe dreams~$\acyclicPipeDreams(126543)$ connected by a sequence of increasing flips in~$\pipeDreams(126543)$ but by no sequence of increasing flips in~$\acyclicPipeDreams(126543)$.}
	\label{fig:counterExampleRestrictionIncreasingFlipPoset}
\end{figure}
\end{remark}

To prove \cref{thm:pipeDreamQuotient}, we need the following auxiliary statement.

@@

\begin{proof}
Say that~$i < j$ while~$\omega^{-1}(i) > \omega^{-1}(j)$ and that~$i \to j$ is an arc of~$P\contact$ while~$j \to i$ is an arc of~$P'{}\contact$.
Since $P\contact$ is acyclic, there is no path from~$j$ to~$i$ in~$P\contact$.
Assume by means of contradiction that there is a path~$i \to k_1 \to \dots \to k_p \to j$ in~$P\contact$ with~$p \ge 1$.
Since the arcs of~$P'{}\contact$ are the arcs of~$P\contact$ where only extremities~$i$ and~$j$ can be changed, $P'{}\contact$ contains the path~$k_1 \to \dots \to k_p$ and at least one of the arcs~$i \to k_1$ or~$j \to k_1$, and at least one of the arcs~$k_p \to j$ or~$k_p \to i$.
Consequently, since~$P'{}\contact$ contains the arc~$j \to i$ and is acyclic, it must contain the path~$j \to k_1 \to \dots \to k_p \to i$.
We thus obtained that~$i \contactLess{P} k_1 \contactLess{P} j$ while~$i \contactMore{P'} k_1 \contactMore{P'} j$, and~$k_1$ has a contact with~$i$ in~$P$ that becomes a contact with~$j$ in~$P'$.

Consider now the contact~$c$ of~$P$ which is a crossing in~$P'$ and the contact~$c'$ of~$P'$ which is a crossing of~$P$.
Let~$R$ the rectangle with corners~$c$ and~$c'$.
Since~$k_1$ has a contact with~$i$ in~$P$ and with~$j$ in~$P'$, it must pass inside~$R$.
Since~$i \contactLess{P} k_1 \contactLess{P} j$ and~$i \contactMore{P'} k_1 \contactMore{P'} j$, the pipe~$k$ has no elbow located northwest or southeast of~$c$ or~$c'$, hence no elbow located north, south, west or east of~$R$.
We thus obtain that~$k$ must be straight before it reaches~$R$, and after it leaves~$R$.
Hence $k_1 < i < j$ and~$\omega^{-1}(k_1) > \omega^{-1}(j) > \omega^{-1}(i)$.
By \cref{lem:consequenceRectangle1}, this contradicts~$i \contactLess{P} k_1$~and~$k_1 \contactMore{P'} j$.
\end{proof}

\begin{proof}[Proof of \cref{thm:pipeDreamQuotient}]
We need to prove that the following conditions are equivalent for two distinct pipe dreams~$P,P' \in \acyclicPipeDreams(\omega)$:
\begin{enumerate}[(i)]
\item there is an increasing flip from $P$ to~$P'$,
\item there exist linear extensions~$\pi$ of~$P$ and~$\pi'$ of~$P'$ such that~$\pi'$ covers~$\pi$ in weak order.
\end{enumerate}
\cref{lem:lowerSetPipeDreams,thm:partitionPipeDreams} directly imply that (ii) $\Rightarrow$ (i).
For (i) $\Rightarrow$ (ii), let~$i < j$ be the two pipes involved in the flip between~$P$ and~$P'$.
Hence, $i \to j$ is an arc of~$P\contact$ while~$j \to i$ is an arc of~$P'{}\contact$.
By \cref{lem:flippable}, there is no directed path from~$i$ to~$j$ in~$P\contact$ besides the arcs~$i \to j$ (there might be more than one such arc).
Hence, there exists a linear extension~$\pi$ of~$P$ where~$i$ and~$j$ are consecutive.
Write~$\pi \eqdef UijV$ and define~$\pi' \eqdef UjiV$.
Since~$i \to j$ is an arc of~$P\contact$, $\pi'$ is not a linear extension of~$P$.
Hence, by \cref{lem:lowerSetPipeDreams}, $\pi'$ is a linear extension of~$P'$.
\end{proof}

@@

\begin{proposition}
For any pipe dreams~$P, P' \in \acyclicPipeDreams(\omega)$, the following assertions are equivalent:
\begin{enumerate}[(i)]
\item there is a path from $P$ to~$P'$ in the increasing flip graph on~$\acyclicPipeDreams(\omega)$,
\item there exist linear extensions~$\pi$ of~$P$ and~$\pi'$ of~$P'$ such that~$\pi < \pi'$ in weak order,
\item the minimal (resp.~max) linear extensions~$\pi$ of~$P$ and~$\pi'$ of~$P$ satisfy~$\pi < \pi'$ in weak order.
\item there is no~$i < j$ such that~$i \contactMore{P} j$ and~$i \contactLess{P'} j$,
\item for all~$i < j$, if~$i \contactMore{P} j$, then~$j \contactMore{P'} i$,
\item for all~$i < j$, if~$i \contactLess{P'} j$, then~$i \contactLess{P} j$,
\end{enumerate}
\end{proposition}

\begin{proof}
We already proved the equivalence (i) $\Leftrightarrow$ (ii) in \cref{thm:pipeDreamQuotient}.
The equivalence (ii) $\Leftrightarrow$ (iii) is valid for any lattice quotient.
The equivalences (iii) $\Leftrightarrow$ (iv) $\Leftrightarrow$ (v) $\Leftrightarrow$ (vi) follow from the descriptions of \cref{prop:WOIP} for the inversions of the minimum and the non-inversions of the maximum of a weak order interval.
\end{proof}

@@

\vincent{Explain with the example of the sylvester congruence}

@@

\vincent{I want to think about congruences of the interval $[e,\omega]$ in general.}
